### 2024-12-11 paper-en.tex
Refactoring

The following definitions will be used onwards. We will use the definition of a bipartite graph based on a definition 7.1 given in a book % <<Дискретная Математика>> Я.М. Ерусалимского % Сослаться на книжку Татта (Tutte)

@@

\begin{theorem}[on wings of cycles]\label{theorem_wings}
    Let $\eta = (e_1, ..., e_d)$ and $\mu = (\epsilon_1, ..., \epsilon_{\delta})$ be different paths on metagraph $G$, with first vertex $a$, and last vertex $b$, $\chi(\eta) = \chi(\mu) (\mod{r})$, and let $\gamma = (e_1, \dots, e_d, \epsilon_{\delta}, \dots, \epsilon_1)$ be a path on $G$. Then paths on expansion which correspond to $\gamma$, are cycles.
\end{theorem}

@@

Let $G(V, E, f, w)$ be a metagraph and $w(e_i) = x_i \; \forall i = 0, \dots, |E|$, where $x_i$ are variables that will help us build a graph with given girth.

@@

Let $G(A \cup B, E, f, w)$ be a metagraph and let $w(e_i) = x_i \; \forall i = 0..|E|$.

@@

Assign the system of all inequalities $\mathcal{S}$ to be empty.
For each vertex $a \in A$:

\begin{itemize}
    \item Add starting label $(a, 0, 0, nil)$ to an empty set of all labels,
    \item Iterating by $l$ from $1$ to $g / 2$:
    \begin{itemize}
        \item For all labels $(v, l - 1, ch, p)$:
        \item
            % Что-то тут не чисто с p
            For each edge $e \not= p$ incident to vertex $v$,
            build label $(v', l, ch', p)$, where $ch' = ch + (-1)^{l} w(p)$, $v'$ is a vertex different from $v$, incident to $p$.
            Add this label to the set of all labels.
    \end{itemize}
    \item for all pairs of labels $(v_1, l_1, ch_1, p_1)$ and $(v_2, l_2, ch_2, p_2)$ such that $v_1 = v_2$ and $l_1 = l_2$, add inequality $ch_1 \neq \ch_2$ to the system $\mathcal{S}$.
\end{itemize}

@@

In figure \ref{neq_system_graph} the metagraph $G$ with for edges $e_1, e_2, e_3, e_4$ is shown. Weights of edges are variables $x_1, x_2, x_3, x_4$.

@@

        \bezier{300}(75,165)(10,100)(35,35)
        \put(5,100){$x_1$}

        \bezier{300}(75,165)(56,100)(35,35)
        \put(56,80){$x_2$}

        \bezier{300}(75,165)(140,100)(115,35)
        \put(125,100){$x_3$}

        \bezier{300}(75,165)(94,100)(115,35)
        \put(87,80){$x_4$}
    \end{picture}
    \caption{ Metagraph $G$, with variable weights of edges $x_1, x_2, x_3, x_4$. }
    \label{neq_system_graph}
\end{figure}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{ | c | c | c | }
        \hline
        $g = 0$            & $g = 1$               & $g = 2$                   \\ \hline
        $(0, v_1, 0, nil)$ & $(1, v_2,  x_1, e_1)$ & $(2, v_1,  x_1 - x_2, b)$ \\ \hline
                           & $(1, v_2,  x_2, e_2)$ & $(2, v_1,  x_2 - x_1, a)$ \\ \hline
                           & $(1, v_3,  x_3, e_3)$ & $(2, v_1,  x_3 - x_4, e)$ \\ \hline
                           & $(1, v_3,  x_4, e_4)$ & $(2, v_1,  x_4 - x_3, d)$ \\ \hline
    \end{tabular}
    \caption{ The set of labels constructed during the execution of the Algorithm 1. }
    \label{cycle_search_table_neq}
\end{table}

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            x_1 - x_2 &\neq 0             (\mod{r}) ; \\
            x_3 - x_4 &\neq 0             (\mod{r}) ; \\
            2 x_1 - 2 x_2 &\neq 0         (\mod{r}) ; \\
            2 x_3 - 2 x_4 &\neq 0         (\mod{r}) ; \\
            x_1 - x_2 - x_3 + x_4 &\neq 0 (\mod{r}) ; \\
            x_1 - x_2 + x_3 - x_4 &\neq 0 (\mod{r}) . \\
        \end{array}
        \label{eqs:cycle_search_neqs}
    \right.
\end{equation}

@@

% Абзац про минимизацию r

@@

\begin{enumerate}
    \item Let $e$ be the number of variables in a system of inequalities.
    \item If the system contains inequality $0 \neq 0$ then it has no solution, the algorithm halts.
    \item Select all inequalities of the form $a x_e + b \neq 0$. Denote the set of all such inequalities as $I$.
    \item Let $v$ be a number such that $v \neq -b/a $ for each inequality $"a x_e + b \neq 0" \in I$. In case $I = \varnothing$ assign $v = 0$. Such a number can be found since $I$ is finite.
    \item Substitute value $v$ instead of variable $x_e$ in all inequalities and add equality $x_e = v$ to the solution. By $v$ selection conditions, none of inequalities will turn into $0 \neq 0$ by such a substitution. 
    \item If $e > 1$ assign $e = e - 1$ and return to step 3, otherwise the solution is found, proceed to calculation of parameter $r$.
\end{enumerate}

@@

Assign $x_1 = 0$. Recalculate considered system with assigned value. We get

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - x_2 &\neq 0            ; \\
            x_3 - x_4 &\neq 0        ; \\
            - 2 x_2 &\neq 0          ; \\
            - 2 x_4 &\neq 0          ; \\
            - x_2 - x_3 + x_4 &\neq 0; \\
            - x_2 + x_3 - x_4 &\neq 0. \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_1}
\end{equation}

Variable $x_2$ can not be equal to $0$ since it would violate inequality $- x_2 \neq 0$. Assign $x_2 = 1$. We obtain

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0             ; \\
            x_3 - x_4 &\neq 0       ; \\
            - 2 &\neq 0             ; \\
            - 2 x_4 &\neq 0         ; \\
            - 1 - x_3 + x_4 &\neq 0 ; \\
            - 1 + x_3 - x_4 &\neq 0 . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_2}
\end{equation}

Assigning $x_3 = 0$ similarly, we get

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0       ; \\
            - x_4 &\neq 0     ; \\
            - 2 &\neq 0       ; \\
            - 2 x_4 &\neq 0   ; \\
            - 1 + x_4 &\neq 0 ; \\
            - 1 - x_4 &\neq 0 . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_3}
\end{equation}

Variable $x_4$ can not be equal to $0, 1$ and $-1$ since they violate the last inequalities. Assign $x_4 = 2$. We finally obtain

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0   ; \\
            - 2 &\neq 0   ; \\
            - 2 &\neq 0   ; \\
            - 4 &\neq 0   ; \\
              1 &\neq 0   ; \\
            - 3 &\neq 0   . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_4}
\end{equation}

% TODO: пронумеровать примеры и вставить нормальную ссылку

This system does not allow $r$ to be $1, 2, 3$ and $4$, thus the least possible value of $r$ is $5$. 5-expansion of metagraph $G$ considered in example $<ref>$ with assigned weights, is shown in figure \ref{neq_system_res}. Girth of this graph is equal to $6$, and one of the cycles with length $6$ is marked in bold.

@@

\begin{notice}
    For any metagraph that has $K_{2,3}$ as a subgraph, the girth of any its r-expansion is at most 12.
    Consider metagraph $G$ shown in figure \ref{full_graph_2_by_3}. Let $\mu = (e_1, e_4, e_5, e_2, e_3, e_6, e_4, e_1, e_2, e_5, e_6, e_3)$ be a cycle. Its characteristic is euqal to $x_1 - x_4 + x_5 - x_2 + x_3 - x_6 + x_4 - x_1 + x_2 - x_5 + x_6 - x_3 = 0$. Therefore, paths that correspond to $\mu$ are cycles with length of $12$ on r-expansions of $G$.
\end{notice}

@@

    1. the first and the last vertices $a$ and $b$ of path $\eta$ (possibly equal) belong to cycle $\mu$,

@@

\textbf{Теорема.}

@@

Consider set of cycles $C$ on metagraph $G'$ with length less then $2 \theta$ such that for each cycle $c \in C$ there exists an expansions of $G'$ on which $c$ corresponds to non-cycles only.

We can use Algorithms 1 and 2 to select weights $w$ and parameter $r$ such that all cycles from $C$ correspond to non-cycles on all expansions determined by pairs $(w, r)$. Thus we can consider only expansions, where cycles from $C$ are broken. Therefore, consider only cycles on metagraph which correspond to cycles on all expansions. Note that all expansions determined by pairs $(w, r)$ have similar girth. Therefore. we can consider the shortest of such cycles only.

@@

% Заключение: 
% <"повторить" последний абзац введения>
% Планы на будущее: обобщить концепцию метаграфов для произвольных (не двудольных) графов

@@

@@

%% G. A. MARGULIS EXPLICIT CONSTRUCTIONS OF GRAPHS WITHOUT SHORT CYCLES AND LOW DENSITY CODES 
%% https://sci-hub.se/10.1007/bf02579283

%% Imrich, W. (1984). Explicit construction of regular graphs without small cycles
%% https://sci-hub.se/10.1007/bf02579157

%% David J.C. MacKay, Michael S. Postol, Weaknesses of Margulis and Ramanujan–Margulis Low-Density Parity-Check Codes
%% https://www.inference.org.uk/mackay/margulis.pdf
%% Критика метода Маргулиса

>>>>

The following definitions will be used onwards. We will use the definition of a bipartite graph based on a definition 1 given in a book \cite{tutte_graph_definition}.

@@

\begin{theorem}[on wings of cycles]\label{theorem_wings}
    Let $\eta = (e_1, ..., e_d)$ and $\mu = (\epsilon_1, ..., \epsilon_{\delta})$ be different paths on metagraph $G$, with first vertex $x$, and last vertex $y$, $\chi(\eta) = \chi(\mu) (\mod{r})$, and let $\gamma = (e_1, \dots, e_d, \epsilon_{\delta}, \dots, \epsilon_1)$ be a path on $G$. Then paths on expansion which correspond to $\gamma$, are cycles.
\end{theorem}

@@

Let $G(V, E, f, w)$ be a metagraph and $w(e_i) = w_i \; \forall i = 0, \dots, |E|$, where $w_i$ are variables that will help us build a graph with given girth.

@@

Let $G(A \cup B, E, f, w)$ be a metagraph and let $w(e_i) = w_i \; \forall i = 0..|E|$.

@@

Assign the system of all inequalities $\mathcal{S}$ to be empty.
For each vertex $x \in A$:

\begin{itemize}
    \item Add starting label $(x, 0, 0, nil)$ to an empty set of all labels,
    \item Iterating by $l$ from $1$ to $g / 2$:
    \begin{itemize}
        \item For all labels $(v, l - 1, ch, p)$:
        \item
            For each edge $e \not= p$ incident to vertex $v$,
            build label $(v', l, ch', p)$, where $ch' = ch + (-1)^{l} w(p)$, $v'$ is a vertex different from $v$, incident to $p$.
            Add this label to the set of all labels.
    \end{itemize}
    \item for all pairs of labels $(v_1, l_1, ch_1, p_1)$ and $(v_2, l_2, ch_2, p_2)$ such that $v_1 = v_2$ and $l_1 = l_2$, add inequality $ch_1 \neq \ch_2 (\mod{r})$ to the system $\mathcal{S}$.
\end{itemize}

@@

In figure \ref{neq_system_graph} the metagraph $G$ with for edges $e_1, e_2, e_3, e_4$ is shown. Weights of edges are variables $w_1, w_2, w_3, w_4$.

@@

        \bezier{300}(75,165)(10,100)(35,35)
        \put(5,100){$w_1$}

        \bezier{300}(75,165)(56,100)(35,35)
        \put(56,80){$w_2$}

        \bezier{300}(75,165)(140,100)(115,35)
        \put(125,100){$w_3$}

        \bezier{300}(75,165)(94,100)(115,35)
        \put(87,80){$w_4$}
    \end{picture}
    \caption{ Metagraph $G$, with variable weights of edges $w_1, w_2, w_3, w_4$. }
    \label{neq_system_graph}
\end{figure}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{ | c | c | c | }
        \hline
        $g = 0$            & $g = 1$               & $g = 2$                   \\ \hline
        $(0, v_1, 0, nil)$ & $(1, v_2,  w_1, e_1)$ & $(2, v_1,  w_1 - w_2, b)$ \\ \hline
                           & $(1, v_2,  w_2, e_2)$ & $(2, v_1,  w_2 - w_1, a)$ \\ \hline
                           & $(1, v_3,  w_3, e_3)$ & $(2, v_1,  w_3 - w_4, e)$ \\ \hline
                           & $(1, v_3,  w_4, e_4)$ & $(2, v_1,  w_4 - w_3, d)$ \\ \hline
    \end{tabular}
    \caption{ The set of labels constructed during the execution of the Algorithm 1. }
    \label{cycle_search_table_neq}
\end{table}

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            w_1 - w_2 &\neq 0             (\mod{r}) ; \\
            w_3 - w_4 &\neq 0             (\mod{r}) ; \\
            2 w_1 - 2 w_2 &\neq 0         (\mod{r}) ; \\
            2 w_3 - 2 w_4 &\neq 0         (\mod{r}) ; \\
            w_1 - w_2 - w_3 + w_4 &\neq 0 (\mod{r}) ; \\
            w_1 - w_2 + w_3 - w_4 &\neq 0 (\mod{r}) . \\
        \end{array}
        \label{eqs:cycle_search_neqs}
    \right.
\end{equation}

@@

@@

\begin{enumerate}
    \item Let $e$ be the number of variables in a system of inequalities.
    \item If the system contains inequality $0 \neq 0$ then it has no solution, the algorithm halts.
    \item Select all inequalities of the form $a x_e + b \neq 0$. Denote the set of all such inequalities as $I$.
    \item Let $v$ be a number such that $v \neq -b/a $ for each inequality $"a x_e + b \neq 0 (\mod{r})" \in I$. In case $I = \varnothing$ assign $v = 0$. Such a number can be found since $I$ is finite.
    \item Substitute value $v$ instead of variable $x_e$ in all inequalities and add equality $x_e = v$ to the solution. By $v$ selection conditions, none of inequalities will turn into $0 \neq 0$ by such a substitution. 
    \item If $e > 1$ assign $e = e - 1$ and return to step 3, otherwise the solution is found, proceed to calculation of parameter $r$.
\end{enumerate}

@@

Assign $w_1 = 0$. Recalculate considered system with assigned value. We get

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - w_2 &\neq 0             (\mod{r}); \\
            w_3 - w_4 &\neq 0         (\mod{r}); \\
            - 2 w_2 &\neq 0           (\mod{r}); \\
            - 2 w_4 &\neq 0           (\mod{r}); \\
            - w_2 - w_3 + w_4 &\neq 0 (\mod{r}); \\
            - w_2 + w_3 - w_4 &\neq 0 (\mod{r}). \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_1}
\end{equation}

Variable $w_2$ can not be equal to $0$ since it would violate inequality $- w_2 \neq 0$. Assign $w_2 = 1$. We obtain

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0              (\mod{r}); \\
            w_3 - w_4 &\neq 0        (\mod{r}); \\
            - 2 &\neq 0              (\mod{r}); \\
            - 2 w_4 &\neq 0          (\mod{r}); \\
            - 1 - w_3 + w_4 &\neq 0  (\mod{r}); \\
            - 1 + w_3 - w_4 &\neq 0  (\mod{r}). \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_2}
\end{equation}

Assigning $w_3 = 0$ similarly, we get

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0        (\mod{r}); \\
            - w_4 &\neq 0      (\mod{r}); \\
            - 2 &\neq 0        (\mod{r}); \\
            - 2 w_4 &\neq 0    (\mod{r}); \\
            - 1 + w_4 &\neq 0  (\mod{r}); \\
            - 1 - w_4 &\neq 0  (\mod{r}). \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_3}
\end{equation}

Variable $w_4$ can not be equal to $0, 1$ and $-1$ since they violate the last inequalities. Assign $w_4 = 2$. We finally obtain

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0    (\mod{r}) ; \\
            - 2 &\neq 0    (\mod{r}) ; \\
            - 2 &\neq 0    (\mod{r}) ; \\
            - 4 &\neq 0    (\mod{r}) ; \\
              1 &\neq 0    (\mod{r}) ; \\
            - 3 &\neq 0    (\mod{r}) . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs_4}
\end{equation}

This system does not allow $r$ to be $1, 2, 3$ and $4$, thus the least possible value of $r$ is $5$. 5-expansion of metagraph $G$ considered in figure \ref{neq_system_graph} with assigned weights, is shown in figure \ref{neq_system_res}. Girth of this graph is equal to $6$, and one of the cycles with length $6$ is marked in bold.

@@

\begin{notice}
    For any metagraph that has $K_{2,3}$ as a subgraph, the girth of any its r-expansion is at most 12.
    Consider metagraph $G$ shown in figure \ref{full_graph_2_by_3}. Let $\mu = (e_1, e_4, e_5, e_2, e_3, e_6, e_4, e_1, e_2, e_5, e_6, e_3)$ be a cycle. Its characteristic is euqal to $w_1 - w_4 + w_5 - w_2 + w_3 - w_6 + w_4 - w_1 + w_2 - w_5 + w_6 - w_3 = 0$. Therefore, paths that correspond to $\mu$ are cycles with length of $12$ on r-expansions of $G$.
\end{notice}

@@

    1. the first and the last vertices $x$ and $y$ of path $\eta$ (possibly equal) belong to cycle $\mu$,

@@

@@

Consider set of cycles $\mathcal{H}_G$ on metagraph $G'$ with length less then $2 \theta$ such that for each cycle $c \in C$ there exists an expansions of $G'$ on which $c$ corresponds to non-cycles only.

We can use Algorithms 1 and 2 to select weights $w$ and parameter $r$ such that all cycles from $\mathcal{H}_G$ correspond to non-cycles on all expansions determined by pairs $(w, r)$. Thus we can consider only expansions, where cycles from $\mathcal{H}_G$ are broken. Therefore, consider only cycles on metagraph which correspond to cycles on all expansions. Note that all expansions determined by pairs $(w, r)$ have similar girth. Therefore. we can consider the shortest of such cycles only.

@@

@@

\begin{thebibliography}{0}

@@

\bibitem{margulis_construction}
G. A. MARGULIS EXPLICIT CONSTRUCTIONS OF GRAPHS WITHOUT SHORT CYCLES AND LOW DENSITY CODES 
%% https://sci-hub.se/10.1007/bf02579283

\bibitem{imrich_construction}
Imrich, W. (1984). Explicit construction of regular graphs without small cycles
%% https://sci-hub.se/10.1007/bf02579157

%% Критика метода Маргулиса
\bibitem{margulis_critique}
David J.C. MacKay, Michael S. Postol, Weaknesses of Margulis and Ramanujan–Margulis Low-Density Parity-Check Codes
% https://www.inference.org.uk/mackay/margulis.pdf

%% Discard

\end{thebibliography}
